Characterizing Exposure to and Sharing Knowledge of Drivers of Environmental Change in the St. Lawrence System

### 2020-04-24 Revisions2/Beauchesne-eDriversMS.Rmd
finish answer to reviewer comments

---
title: Next Generation Planning -- Characterizing Exposure to and Shared Knowledge on Drivers of Ecological Change in the St. Lawrence System
author: David Beauchesne^1,2,\*^, Rémi M. Daigle^2^, Steve Vissault^3^, Dominique Gravel^3^, Andréane Bastien^4^, Simon Bélanger^5^, Pascal Bernatchez^5^, Marjolaine Blais ^6^, Hugo Bourdages^6^, Clément Chion^7^, Peter S. Galbraith^6^, Benjamin Halpern^8,9^, Camille Lavoie^2^, Christopher W. McKindsey^6^, Alfonso Mucci^10^, Simon Pineault^11^, Michel Starr^6^, Anne-Sophie Ste-Marie^4^, Philippe Archambault^2^
fontsize: 12pt
font-family: Times New Roman
output:
  pdf_document:
    toc: false
    number_sections: true
    keep_tex: true
  word_document:
    toc: false
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{listings}
   - \usepackage[svgnames]{xcolor}
   - \usepackage{lscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
   - \usepackage{bbm}
   - \usepackage{float}
   - \usepackage[normalem]{ulem}
   - \definecolor{blue}{HTML}{20cad5}
relativeurls: true
bibliography: Beauchesne-eDriversMS.bib
csl: frontiers.csl
link-citations: yes
---

@@

<!--
To discuss with co-authors

- New article title
- Should we present how the platform is used in the analyses or really keep
  that portion fully separate? If we include it, modify answer to question 2
  comment 1

-->

@@

The St. Lawrence is a vast and complex socio-ecological system providing a
wealth of services sustaining numerous economic sectors. ~~\textcolor{red}{These ecosystems are}~~
\textcolor{blue}{This ecosystem is}
subject to significant human pressures that overlap and potentially interact
with climate driven environmental changes. Our objective in this paper is to
systematically characterize the distribution and intensity of drivers \textcolor{blue}{of change} in the
St. Lawrence System. ~~To do so, we launch *eDrivers*, an open knowledge
platform gathering experts committed to structuring, standardizing and sharing
knowledge on drivers \textcolor{blue}{of change} in support of science and management.~~ We gathered data on
22 coastal, climate, fisheries and marine traffic drivers through
collaborations, existing environmental initiatives and open data portals. We
show that few areas of the St. Lawrence are free of cumulative exposure. The
Estuary, ~~the~~ Anticosti Gyre and coastal areas are particularly exposed,
especially in the vicinity of urban centers. We identified 6 areas of distinct
cumulative exposure regime that show that certain drivers \textcolor{blue}{of change} typically co-occur
in different regions of the St. Lawrence and that coastal areas are exposed to
all driver types. Of particular concern are two threat complexes capturing
most exposure hotspots that show the convergence of contrasting exposure
regimes at the head of the Laurentian Channel.
\textcolor{blue}{Sharing the knowledge acquired emerged as a priority to minimize future
environmental assessment efforts. We thus launch *eDrivers*, an open knowledge
platform gathering experts committed to structuring, standardizing and sharing
knowledge on drivers of change in support of science and management.}
*eDrivers* was built on a
series of guiding principles upholding existing data management and open
science standards. We therefore expect it to evolve through time to address
knowledge gaps and refine current driver layers. Ultimately, we believe that
*eDrivers* represents a much needed solution that could radically influence
broad scale research and management practices by increasing knowledge accessibility
and interoperability.

**Keywords**: ocean observing systems, St. Lawrence, environmental drivers,
cumulative exposure, threat complex, multiple stressors, global change

@@

The St. Lawrence System, formed by one of the largest estuaries in the world\textcolor{red}{\sout{, the St. Lawrence Estuary,}}
and a vast interior sea\textcolor{red}{\sout{, the Gulf of St. Lawrence}}, is
a complex social-ecological system characterized by highly variable
environmental conditions and oceanographic processes
[@el-sabh1990a; @white1997; @dufour2007]. It constitutes a unique
and heterogeneous array of habitats suited for the establishment of diverse and
productive ecological communities [@savenkoff2000]. As a result, the St.
Lawrence System provides a wealth of ecosystem services that have \textcolor{red}{\sout{historically and contemporarily}}
benefited the Canadian economy. It sustains a rich fisheries
industry targeting more than 50 species, serves as the gateway to eastern
North-America by granting access to more than 40 ports and the most densely
populated Canadian region, hosts a booming tourism industry and an expanding
aquaculture production, fosters emerging activities and boasts a yet untapped
hydrocarbon potential [@beauchesne2016; @schloss2017; @archambault2017]. With
major investments recently made and more forthcoming in economic and
infrastructure development and research [*e.g.* @governmentofquebec2015; @rqm2018], an intensification of the human footprint is likely in the St. Lawrence System.

As elsewhere in the world [see @halpern2015a; @halpern2019], this intensifying human footprint
will likely result in increasingly intricate environmental exposure regimes,
*i.e.* suites of overlapping \textcolor{red}{\sout{and potentially interacting}} environmental drivers \textcolor{blue}{of change}
threatening ecosystems, habitats or ecological communities. Drivers \textcolor{blue}{of change}, often referred to as stressors or pressures \textcolor{blue}{and hereafter called drivers for simplicity}, are any externalities that affect
environmental processes and disturb natural systems. Drivers may originate from
natural or human-induced biophysical processes (*e.g.* sea surface\textcolor{blue}{-water} temperature
anomalies and hypoxia) or directly from anthropogenic activities (*e.g.*
fisheries and marine pollution). The potential for complex interactions between
\textcolor{blue}{co-occurring} drivers is the largest uncertainty when studying or predicting environmental impacts [@darling2008; @cote2016]. Multiple drivers can combine non-linearly and result in effects that are greater (synergistic effect) or lower (antagonistic effect) than the sum of individual effects
[@crain2008; @darling2008; @cote2016].

The uncertainty associated with complex driver interactions must therefore be taken into account when investigating environmental impacts [@cote2016], yet most research on driver effects in marine environments remains overwhelmingly focused on single driver assessments [@obrien2019]. Increasing exposure and the experiences of past ecological tragedies such as the collapse of cod fisheries [@frank2005; @dempsey2018] and the decline of the beluga and right whale populations [@plourde2014] together urge the need to characterize the distribution, intensity and \textcolor{red}{\sout{overlap}} \textcolor{blue}{co-occurrence} between drivers in the St. Lawrence System. \textcolor{blue}{The impacts of drivers, whether individually or in combination, depend on species-specific vulnerabilities. Driver co-occurrence alone is therefore not sufficient evidence to claim interactions among drivers. However, as drivers must co-occur to interact, the likelihood of interactions will increase with the number of drivers co-occurring.} \textcolor{red}{\sout{This will}} \textcolor{blue}{Describing drivers will therefore} provide critical information on areas most exposed to \textcolor{red}{\sout{cumulative}} drivers \textcolor{red}{\sout{and on the interaction potential of drivers}} in the St. Lawrence. It is also a necessary step towards the holistic and integrated management of the St. Lawrence System.

\textcolor{blue}{Holistic approaches typically involve, but are not limited to, the following steps: selecting and describing valued ecosystem components (\textit{e.g.} habitats and species) and drivers (\textit{e.g.} marine traffic and ocean acidification), assessing the exposure and vulnerability of valued components to drivers, selecting a proper spatio-temporal scale, monitoring, and public and stakeholder participation} [@dube2001].
Gathering environmental
\textcolor{red}{\sout{data for large scale, systematic}}
\textcolor{blue}{knowledge for holistic}
initiatives can, however, be a very challenging and time consuming – not to say painful – process. On one hand, there is an overwhelming and expanding wealth of data available. Such information overload may inhibit our ability to make decisions based on scientific information, promote massive effort duplication, disproportionately appropriate research funds to certain sectors, and obscure knowledge gaps amid a sea of information [@eppler2004]. On the other hand, crucial data are lacking and remain largely unavailable or inaccessible for a variety of reasons, including proprietary rights, lack of organizational time, capacity and training, and in some \textcolor{red}{\sout{rare}} cases unwillingness to share, curtailing our ability for appropriate decision-making.

There are now initiatives that \textcolor{blue}{facilitate the data gathering process} \textcolor{red}{\sout{address this issue}} by assembling, organizing and sharing environmental knowledge, such as the Ocean Biogeographic Information
System [OBIS; @obis2019] for biotic data and Bio-ORACLE [@tyberghein2012] for
abiotic data.
However, equivalent platforms for environmental drivers have largely focused on
single drivers (e.g. Global Fishing Watch) and platforms collating
data and knowledge on multiple drivers in a comparable and interoperable way
remain conspicuously missing [but see @halpern2015]. This is in spite of
integrated management and assessment approaches needing efficient data reporting,
standardized data management practices and tools tailored to the study of the
effects of multiple drivers [@dafforn2016; @stock2018].
~~An additional objective
thus emerged in the process of addressing our initial goals:
sharing information about the
distribution and intensity of drivers of change in the St.Lawrence.~~

Our main ~~objective~~ \textcolor{blue}{goal} in this study is to characterize the distribution and
intensity of drivers in the St. Lawrence. More specifically, ~~we aim~~ \textcolor{blue}{our objectives are} to: 1)
identify areas of high cumulative exposure\textcolor{red}{\sout{, 2) identify drivers that are likely to interact in the St. Lawrence}} and \textcolor{red}{\sout{3)}} \textcolor{blue}{2)} characterize areas with similar cumulative exposure regimes.
\textcolor{red}{\sout{Here we focus solely on drivers, one of four elements required to evaluate environmental impacts, *i.e.* drivers, ecosystem components of interest (*e.g.* habitats or species) and the exposure and the vulnerability of the ecosystem components to drivers.}}
\textcolor{blue}{Here, we focus solely on drivers, aiming at better understanding their distribution in the St. Lawrence and at facilitating the widespread availability of their characterization for holistic assessments and management approaches.}
\textcolor{blue}{An additional objective emerged while addressing the main objective of this
manuscript: sharing information about the distribution and intensity of drivers of change in the St.Lawrence.}
We achieve ~~these objectives~~ \textcolor{blue}{this} with the development of an
open knowledge platform,~~named~~ *eDrivers*,~~. The platform was~~ designed to
facilitate collaboration, real-time assessments of cumulative exposure and to
evolve ~~with the addition of information and threats to~~
\textcolor{blue}{as knowledge accumulates and new threats emerge in}
the St-Lawrence ecosystems.
\textcolor{blue}{Here, we present its guidng principles and accompanying tools.}

@@

The St. Lawrence drains over 25% of global freshwater reserves through its
connexion to the Great Lakes Basin\textcolor{red}{\sout{. The Basin}}\textcolor{blue}{, which }is home to over 45 million
North-Americans, *i.e.* 15 and 30 million in Canada and the United States,
respectively. The St. Lawrence itself boasts a much lower
population of approximately 1 million Canadians living within 10 km of the
coast, with populations mainly located in a few coastal cities in the Estuary
and the Southern Gulf [@statistics-canada2017].

@@

\textcolor{red}{\sout{The list of drivers for which we sought data was informed by}} \textcolor{blue}{Drivers selection was informed by} a global
cumulative impact assessment initiative [@halpern2008a; @halpern2015a; @halpern2019] \textcolor{blue}{and available from the National Center for Ecological Analysis and Synthesis (NCEAS) online data repository (Halpern et al. 2015)},
regional holistic evaluations of the state of the St. Lawrence [@dufour2007;
@benoit2012], and communications with regional experts (Table 1).
\textcolor{red}{\sout{We integrated datasets from regional experts and also use global data from the global cumulative impact assessment initiative [@halpern2008a; @halpern2015a]  available from the National Center for Ecological Analysis and Synthesis (NCEAS) online data repository [Table 1; @halpern2015].}}
~~We selected global data that were
unavailable at the regional scale and that were available at a resolution
adequate for use at the scale of the St. Lawrence (*e.g.* marine pollution).~~
\textcolor{blue}{Where regional data on drivers were unavailable, available global data at a resolution adequate for the scale of the St. Lawrence System were used instead (*e.g.* marine pollution).}

@@

~~Drivers with non-normal frequency distributions were log-transformed, as in
Halpern et al. 2008, to minimize the impact of extreme outliers that may or may
not be real observations (Figure S1), and all drivers were scaled between 0 and
1 to allow driver comparisons. The 99th quantile of individual driver
distribution was used as the upper bound for scaling.~~
\textcolor{blue}{As in Halpern et al. 2019, drivers with non-normal frequency distributions were log-transformed
to avoid underestimating intermediate driver values. All drivers were also scaled
between 0 and 1 to allow driver comparisons. The 99th quantile of individual
driver distribution was used as the upper bound for scaling to control for
extreme values that may or may not be real observations.}
~~All drivers were embedded in a regular grid composed of 245604 $1 km^2$
hexagonal cells to construct the integrated dataset used for the analyses
(Figure S2).~~
\textcolor{blue}{The St. Lawrence System was divided into a regular grid of
1 $km^2$ hexagonal cells in which all drivers were integrated (Figure S2)}

## \textcolor{red}{\sout{Driver interactions}} \textcolor{blue}{Driver co-occurrence}

\textcolor{red}{\sout{Interactions among drivers were investigated using the cumulative footprint ($F$) between pairs of drivers (objective 2). The intensity at which pairs of drivers co-occur was evaluated using a two-dimensional kernel density estimate. As there are 231 pairwise combinations between 22 drivers, we focus on a single example using hypoxia and demersal destructive fisheries, two drivers known to occur mainly in deeper areas of the St. Lawrence and, hence, an interaction between the effects of the two drivers could be anticipated.}}
\textcolor{blue}{As an introduction to indicators of cumulative exposure (objective 1), we begin by presenting a simplifed 2-driver example focusing on hypoxia and demersal destructive fisheries, two drivers known to occur mainly in deeper waters of the St. Lawrence. Driver co-ccurrence was evaluated spatially by summing the scaled intensity of drivers in each grid cell. The intensity at which pairs of drivers co-occur was evaluated using a two-dimensional kernel density.}

@@

Areas with high cumulative exposure (objective 1) were identified by comparing
areas on the basis of the number and relative intensity of drivers in each
grid cell. Throughout the text, we use exposure when describing areas
exposed to drivers and we use footprint when describing the distribution and
intensity of drivers.

Cumulative footprint ($F$) was defined as the sum of the scaled intensity of all
drivers in each grid cell:

$$F_x = \sum_{i=1}^{n} D_{i,x}$$

where $x$ is a grid cell, $i$ is a driver and $D$ is the scaled intensity of
driver $i$. The cumulative footprint provides an estimate of the total relative
footprint in each grid cell. A grid cell with a high $F$ value is either
characterized by multiple drivers at low relative intensity, limited drivers at
high relative intensity, or both.

We also identified cumulative hotspots ($H$) to explore the the distribution of
cumulative exposure in the St. Lawrence (objective 1). Cumulative hotspots ($H$)
were defined as the number of drivers in each grid cell with scaled intensity
contained over their respective 80th percentile:

<!-- Indicator function -->
$$H_x = \sum_{i=1}^{n} \mathbbm{1} (D_{i,x} \; \epsilon \; P_{80, D_i})$$

where, $x$ is a grid cell, $i$ is a driver and $D$ is the scaled intensity of
driver $i$ and $P_{80, D_i}$ is the 80th percentile of driver $i$. Hotspots
thus identify areas where drivers are co-occurring at high relative intensities.

## Threat complexes

In order to \textcolor{red}{\sout{better capture potential interactions between drivers (objective 2) and to}}identify areas with similar exposure regimes (objective \textcolor{red}{\sout{3}}\textcolor{blue}{2}), we identify threat complexes using a clustering approach [*e.g.* see @bowler2019]. We use the term clusters in presenting the methods, but use threat complex when discussing the results on cumulative exposure regimes.

@@

Threat complexes were identified using a partional *k-medoids* clustering
algorithm, CLARA [CLustering for Large Applications; @kaufman1990], which was
designed for large datasets. The CLARA algorithm uses the PAM (Partition
Around Medoids) algorithm on a sample from the original dataset to identify a
set of $k$ objects that are representative of all other objects, *i.e.*
medoids and that are central to the cluster they represent. The goal of the
algorithm is to iteratively minimize intra-cluster dissimilarity. Iterations
are compared on the basis of the average dissimilarity between cluster objects
and representative medoid to select the optimal set of $k$ medoids that
minimizes average dissimilarity. We used the clustering algorithm with the
Manhattan distance since this measure is less affected by extreme values
[@legendre2012], as is the *k-medoids* clustering algorithm [@kaufman1990].
We used 100 iterations using samples of 10000 observations
(*i.e.* ~5% of observations) to identify clusters. Analyses were performed
using the *cluster* R package [@maechler2018].
Partitional clustering algorithms require a user-defined number of clusters.
Values of $k$ ranging from 2 to 10 were tested and validated by selecting
the number of clusters that maximized the average silhouette width [@kaufman1990]
and minimized the total within-cluster sum of squares (Figure S4).

@@

## \textcolor{red}{\sout{Driver interactions}} \textcolor{blue}{Driver co-occurrence}

Hypoxia is mainly distributed in the Laurentian, Anticosti and Esquiman Channels, with the head of the Channels most exposed (Figure \ref{kernel}A). Demersal destructive fisheries are located along the Laurentian Channel, the heads of the Anticosti and Esquiman Channels and around the Magdalen Islands (Figure \ref{kernel}B). By combining both drivers, we can observe that hypoxia and demersal destructive fisheries co-occur mostly at high relative intensity (Figure \ref{kernel}D) in the vicinity of the Anticosti Gyre and the heads of the Esquiman and Anticosti Channels (Figure \ref{kernel}C~~, Box 1~~).

>>>>

---
title: Characterizing Exposure to and Sharing Knowledge of Drivers of Environmental Change in the St. Lawrence System
author: David Beauchesne^1,2,\*^, Rémi M. Daigle^2^, Steve Vissault^3^, Dominique Gravel^3^, Andréane Bastien^4^, Simon Bélanger^5^, Pascal Bernatchez^5^, Marjolaine Blais ^6^, Hugo Bourdages^6^, Clément Chion^7^, Peter S. Galbraith^6^, Benjamin Halpern^8,9^, Camille Lavoie^2^, Christopher W. McKindsey^6^, Alfonso Mucci^10^, Simon Pineault^11^, Michel Starr^6^, Anne-Sophie Ste-Marie^4^, Philippe Archambault^2^
fontsize: 12pt
font-family: Times New Roman
output:
  pdf_document:
    toc: false
    number_sections: true
    keep_tex: true
  word_document:
    toc: false
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{listings}
   - \usepackage[svgnames]{xcolor}
   - \usepackage{lscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
   - \usepackage{bbm}
   - \usepackage{float}
   - \usepackage[normalem]{ulem}
   - \definecolor{blue}{HTML}{20cad5}
relativeurls: true
bibliography: Beauchesne-eDriversMS.bib
csl: frontiers.csl
link-citations: yes
---

@@

@@

The St. Lawrence is a vast and complex socio-ecological system providing a
wealth of services sustaining numerous economic sectors. ~~These ecosystems are~~ \textcolor{blue}{This ecosystem is} subject to significant human pressures that overlap and potentially interact with climate driven environmental changes. Our objective in this paper is to systematically characterize the distribution and intensity of drivers of \textcolor{blue}{environmental} change \textcolor{blue}{(hereafter drivers)} in the St. Lawrence System. ~~To do so, we launch *eDrivers*, an open knowledge platform gathering experts committed to structuring, standardizing and sharing knowledge on drivers of change in support of science and management.~~ We gathered data\textcolor{blue}{-based indicators for} ~~on~~ 22 coastal, climate, fisheries and marine traffic drivers through collaborations, existing environmental initiatives and open data portals. We show that few areas of the St. Lawrence are free of cumulative exposure. The Estuary, ~~the~~ Anticosti Gyre and coastal areas are particularly exposed, especially in the vicinity of urban centers. We identified 6 ~~areas of~~ distinct ~~cumulative exposure regimes~~ \textcolor{blue}{clusters with similar suites of co-occurring drivers} and show that certain driver \textcolor{blue}{combinations are typical of} ~~typically co-occur in~~
different regions of the St. Lawrence and that coastal areas are exposed to
all driver types. Of particular concern are two ~~threat complexes~~\textcolor{blue}{clusters} capturing
most exposure hotspots \textcolor{blue}{and} that show the convergence of contrasting ~~exposure
regimes~~\textcolor{blue}{cumulative exposure profiles} at the head of the Laurentian Channel.
\textcolor{blue}{Sharing the knowledge of drivers acquired emerged as a priority to minimize future environmental assessment efforts. We thus launch \textit{eDrivers}, an open knowledge platform gathering experts committed to structuring, standardizing and sharing knowledge on drivers of environmental change in support of holistic science and management.} *eDrivers* was built on a series of guiding principles upholding existing data management and open science standards. We therefore expect it to evolve through time to address knowledge gaps and refine current driver layers. Ultimately, we believe that *eDrivers* represents a much needed solution that could radically influence broad scale research and management practices by increasing knowledge accessibility and interoperability.

**Keywords**: ocean observing systems, St. Lawrence,
cumulative exposure, multiple stressors, global change

@@

The St. Lawrence System, formed by one of the largest estuaries in the world
and a vast interior sea, is a complex social-ecological system characterized by highly variable environmental conditions and oceanographic processes
[@el-sabh1990a; @white1997; @dufour2007]. It constitutes a unique
and heterogeneous array of habitats suited for the establishment of diverse and
productive ecological communities [@savenkoff2000]. As a result, the St.
benefited the Canadian economy. It sustains a rich fisheries
industry targeting more than 50 species, serves as the gateway to eastern
North-America by granting access to more than 40 ports and the most densely
populated Canadian region, hosts a booming tourism industry and an expanding
aquaculture production, fosters emerging activities and boasts a yet untapped
hydrocarbon potential [@beauchesne2016; @schloss2017; @archambault2017]. With
major investments recently made and more forthcoming in economic and
infrastructure development and research [*e.g.* @governmentofquebec2015; @rqm2018], an intensification of the human footprint is likely in the St. Lawrence System.

~~As elsewhere in the world [see @halpern2015a; @halpern2019], this intensifying human footprint will likely result in increasingly intricate environmental exposure regimes, *i.e.* suites of overlapping drivers of environmental change threatening ecosystems, habitats and ecological communities.~~
\textcolor{blue}{As a result, the St. Lawrence System is exposed to an increasing number of drivers of environmental change, as is observed across ecosystems globally} [see @halpern2015a; @halpern2019].
~~Drivers of environmental change, often referred to as stressors or pressures and hereafter called drivers for simplicity, are any externalities that affect
environmental processes and disturb natural systems.~~
\textcolor{blue}{We broadly define drivers of environmental change (hereafter drivers) as any externality that affects environmental processes and disturbs natural systems.}
Drivers may originate from natural or human-induced biophysical processes (*e.g.* sea surface-water temperature anomalies and hypoxia) or directly from anthropogenic activities (*e.g.* fisheries and marine pollution). The potential for complex interactions between co-occurring drivers is the largest uncertainty when studying or predicting environmental impacts [@darling2008; @cote2016]. Multiple drivers can combine non-linearly and result in effects that are greater (synergistic effect) or lower (antagonistic effect) than the sum of individual effects [@crain2008; @darling2008; @cote2016].

~~The uncertainty associated with complex driver interactions must therefore be taken into account when investigating environmental impacts [@cote2016], yet research on the effects of drivers in marine environments remains overwhelmingly focused on single driver assessments} [@obrien2019].~~  Increasing exposure and the experiences of past ecological tragedies such as the collapse of cod fisheries [@frank2005; @dempsey2018] and the decline of the beluga and right whale populations [@plourde2014] together urge the need to characterize the distribution, intensity and co-occurrence of drivers in the St. Lawrence System. \textcolor{blue}{Research on the effects of drivers in marine environments nonetheless remains overwhelmingly focused on single driver assessments} [@obrien2019]. \textcolor{blue}{While co-occurring drivers may not interact, driver co-occurrence is a requirement for interactions to exist. Knowledge of their co-distribution can therefore identify areas where driver interactions are most likely observed.}
~~Describing drivers will therefore provide critical information on areas most exposed to drivers in the St. Lawrence. It~~

\textcolor{blue}{Characterizing drivers is also a necessary step for the application of holistic management approaches.} Holistic approaches typically involve, but are not limited to, selecting and describing valued ecosystem components (\textit{e.g.} habitats and species) and drivers (\textit{e.g.} marine traffic and ocean acidification), assessing the exposure and vulnerability of valued components to drivers, selecting a proper spatio-temporal scale, monitoring, and public and stakeholder participation [@dube2001].
Gathering environmental knowledge for holistic initiatives can, however, be a very challenging and time consuming –- not to say painful –- process. On one hand, there is an overwhelming and expanding wealth of data available. Such information overload may inhibit our ability to make decisions based on scientific information, promote massive effort duplication, disproportionately appropriate research funds to certain sectors, and obscure knowledge gaps amid a sea of information [@eppler2004]. On the other hand, crucial data are lacking and remain largely unavailable or inaccessible for a variety of reasons, including proprietary rights, lack of organizational time, capacity and training, and in some cases unwillingness to share; this curtails our ability for appropriate decision-making.

There are now initiatives that facilitate the data gathering process by assembling, organizing and sharing environmental knowledge, such as the Ocean Biogeographic Information System [OBIS; @obis2019] for biotic data and Bio-ORACLE [@tyberghein2012] for abiotic data. However, equivalent platforms for ~~environmental~~ drivers have largely focused on single drivers (e.g. Global Fishing Watch) and platforms collating data\textcolor{blue}{-based indicators} and knowledge on multiple drivers in a comparable and interoperable way remain conspicuously missing [but see @halpern2015]. This is in spite of integrated management and assessment approaches needing efficient data reporting, standardized data management practices and tools tailored to the study of the effects of multiple drivers [@dafforn2016; @stock2018]. ~~An additional objective thus emerged in the process of addressing our initial goals: sharing information about the distribution and intensity of drivers of change in the St.Lawrence.~~

Our main ~~objective~~ \textcolor{blue}{goal} in this study is to characterize the distribution and
intensity of drivers in the St. Lawrence. More specifically, ~~we aim~~ \textcolor{blue}{our objectives are} to: 1)
identify areas of high cumulative exposure \textcolor{blue}{to drivers} and 2) characterize areas with similar cumulative exposure ~~regimes~~ \textcolor{blue}{profiles, \textit{i.e.} areas exposed to similar suites of co-occurring drivers}.
\textcolor{blue}{An additional objective emerged while addressing the main goal of this manuscript: sharing information about the distribution and intensity of drivers of environmental change in the St.Lawrence.}
~~We achieve these objectives with the development of an
open knowledge platform, named *eDrivers*. The platform was designed to
facilitate collaboration, real-time assessments of cumulative exposure and to
evolve with the addition of information and threats to the St-Lawrence ecosystems.~~
\textcolor{blue}{We achieve this with the development of an open knowledge platform, \textit{eDrivers}, designed to facilitate the widespread availability of driver characterization for holistic assessments and management approaches. Here, we present its guiding principles and accompanying tools.}

@@

The St. Lawrence drains over 25% of global freshwater reserves through its
connexion to the Great Lakes Basin, which is home to over 45 million
North-Americans, *i.e.* 15 and 30 million in Canada and the United States,
respectively. The St. Lawrence itself boasts a much lower
population of approximately 1 million Canadians living within 10 km of the
coast, with populations mainly located in a few coastal cities in the Estuary
and the Southern Gulf [@statistics-canada2017].

@@

\textcolor{blue}{Drivers, as broadly defined in this study, are data-based indicators of environmental conditions and human activities that are often referred to as driving forces, stressors, pressures or states in the scientific and environmental assessment litterature} [*e.g.* @kristensen2004; @halpern2019].
\textcolor{blue}{Defining such categories, however, can be difficult and is often context- and ecosystem-specific} [@gari2015; @dempsey2018].
\textcolor{blue}{As such, we refrain from articulating our work around a specific framework or imposing categories on data-based products that may change with a user's objective. We rather focus on available data-based indicators that contribute to evaluating an ecosystem's cumulative exposure to multiple threats.}

Drivers selection was informed by a global cumulative impact assessment initiative [@halpern2008a; @halpern2015a; @halpern2019] and available from the National Center for Ecological Analysis and Synthesis (NCEAS) online data repository [@halpern2015], regional holistic evaluations of the state of the St. Lawrence [@dufour2007;
@benoit2012], and communications with regional experts (Table 1).
~~We selected global data that were unavailable at the regional scale and that were available at a resolution adequate for use at the scale of the St. Lawrence (*e.g.* marine pollution).~~ \textcolor{blue}{Where regional data on drivers were unavailable, available global data at a resolution adequate for the scale of the St. Lawrence System were used instead (\textit{e.g.} marine pollution).}

@@

~~Drivers with non-normal frequency distributions were log-transformed, as in
Halpern et al. 2008, to minimize the impact of extreme outliers that may or may
not be real observations (Figure S1), and all drivers were scaled between 0 and
1 to allow driver comparisons. The 99th quantile of individual driver
distribution was used as the upper bound for scaling.~~
\textcolor{blue}{As in Halpern et al. 2019, drivers with non-normal frequency distributions were log-transformed to avoid underestimating intermediate driver values. All drivers were also scaled between 0 and 1 to allow driver comparisons. The 99th quantile of individual driver distribution was used as the upper bound for scaling to control for extreme values that may or may not be real observations.}
~~All drivers were embedded in a regular grid composed of 245604 $1 km^2$
hexagonal cells to construct the integrated dataset used for the analyses
(Figure S2).~~
\textcolor{blue}{The St. Lawrence System was divided into a regular grid of
1 $km^2$ cells in which all drivers were integrated (Figure S2)}

## ~~Driver co-occurrence~~ {-}

~~As an introduction to indicators of cumulative exposure
(objective 1), we begin by presenting a simplifed 2-driver example focusing
on hypoxia and demersal destructive fisheries, two drivers known to occur
mainly in deeper waters of the St. Lawrence. Driver co-ccurrence was evaluated
spatially by summing the scaled intensity of drivers in each grid cell. The
intensity at which pairs of drivers co-occur was evaluated using a two-dimensional kernel density.~~

@@

~~Areas with high cumulative exposure were identified by comparing
areas on the basis of the number and relative intensity of drivers in each
grid cell (objective 1).~~
\textcolor{blue}{We begin by providing a simplifed 2-driver example focusing on
the co-occurrence of hypoxia and demersal destructive fisheries, two drivers
occurring mainly in deeper waters of the St. Lawrence. Driver co-ccurrence was
evaluated spatially by summing the scaled intensity of drivers in each grid cell.
The intensity at which pairs of drivers co-occur was evaluated using a
two-dimensional kernel density. This example demonstrates how driver
co-occurrence was evaluated and serves as a stepping stone to the
integrative indicators used after, \textit{i.e.} cumulative exposure and
cumulative hotspots (objective 1).}

~~Cumulative exposure ($E_C$), was defined as the sum of the scaled intensity of all
drivers in each grid cell:~~
\textcolor{blue}{We evaluated the cumulative exposure ($E_C$) of each grid
cell as the sum of scaled driver intensities:}

$$E_{C_x} = \sum_{i=1}^{n} D_{i,x}$$

where $x$ is a grid cell, $i$ is a driver and $D$ is the scaled intensity of
driver $i$. ~~The cumulative footprint provides an estimate of the total relative
footprint in each grid cell.~~ A grid cell with a high $E_C$ value is either
characterized by multiple drivers at low relative intensity, ~~limited~~ \textcolor{blue}{a few} drivers at
high relative intensity, or both.

~~We also identified cumulative hotspots ($H_C$) to explore the distribution of
cumulative exposure in the St. Lawrence (objective 1). Cumulative hotspots ($H$)
were defined as the number of drivers in each grid cell with scaled intensity
contained over their respective 80th percentile:~~

\textcolor{blue}{We also identified cumulative hotspots ($H_C$) -- \textit{i.e.} areas where drivers co-occur at high
relative intensities -- as the number of drivers in each grid cell with scaled
intensity contained over their respective 80th percentile:}

<!-- Indicator function -->
$$H_{C_x} = \sum_{i=1}^{n} \mathbbm{1} (D_{i,x} \; \epsilon \; P_{80, D_i})$$

where, $x$ is a grid cell, $i$ is a driver and $D$ is the scaled intensity of
driver $i$ and $P_{80, D_i}$ is the 80th percentile of driver $i$.
~~Hotspots thus identify areas where drivers are co-occurring at high relative intensities.~~

## ~~Threat complexes~~\textcolor{blue}{Cumulative exposure profiles}

~~In order to identify areas with similar exposure regimes (objective 2), we identify threat complexes using a clustering approach [*e.g.* see @bowler2019]. We use the term clusters in presenting the methods, but use threat complex when discussing the results on cumulative exposure regimes.~~

@@

\textcolor{blue}{We identified areas with similar cumulative exposure profiles using a clustering approach} [@bowler2019].
~~Threat complexes were identified using~~ \textcolor{blue}{We used} a partional *k-medoids* clustering
algorithm, CLARA [CLustering for Large Applications; @kaufman1990], which was
designed for large datasets. The CLARA algorithm uses the PAM (Partition
Around Medoids) algorithm on a sample from the original dataset to identify a
set of $k$ objects that are representative of all other objects, *i.e.*
medoids and that are central to the cluster they represent. The goal of the
algorithm is to iteratively minimize intra-cluster dissimilarity. Iterations
are compared on the basis of the average dissimilarity between cluster objects
and representative medoid to select the optimal set of $k$ medoids that
minimizes average dissimilarity. We used the clustering algorithm with the
Manhattan distance since this measure is less affected by extreme values
[@legendre2012], as is the *k-medoids* clustering algorithm [@kaufman1990].
We used 100 iterations using samples of 10000 observations
(*i.e.* ~5% of observations) to identify clusters. Analyses were performed
using the *cluster* R package [@maechler2018].
Partitional clustering algorithms require a user-defined number of clusters.
Values of $k$ ranging from 2 to 10 were tested and validated by selecting
the number of clusters that maximized the average silhouette width [@kaufman1990]
and minimized the total within-cluster sum of squares (Figure S4).

@@

## ~~Driver co-occurrence~~ {-}

~~Hypoxia is mainly distributed in the Laurentian, Anticosti and Esquiman Channels,
with the head of the Channels most exposed (Figure \ref{kernel}A). Demersal
destructive fisheries are located along the Laurentian Channel, the heads of the
Anticosti and Esquiman Channels and around the Magdalen Islands (Figure \ref{kernel}B).
By combining both drivers, we can observe that hypoxia and demersal destructive
fisheries co-occur mostly at high relative intensity (Figure \ref{kernel}D) in
the vicinity of the Anticosti Gyre and the heads of the Esquiman and Anticosti
Channels (Figure \ref{kernel}C, Box 1).~~

~~Fisheries in the St. Lawrence have historically affected biodiversity
distribution and habitat quality [@moritz2015]. Concurrently, hypoxia decreases
overall habitat quality, but triggers species-dependent responses ranging from
adaptation [*e.g.* northern shrimp *Pandalus borealis* and Greenland halibut
*Reinhardtius hippoglossoides*; @pillet2016] to reduced growth rates
[@dupont-prinet2013] and avoidance of oxygen-depleted habitats [*e.g.* Atlantic
cod *Gadus morhua*; @chabot2008] to increased mortality [*e.g.* sessile benthic
invertebrates; @eby2005; @gilbert2007; @belley2010]. Certain species may thus be
adversely affected by fisheries and withstand hypoxia but still experience a
decrease in prey availability, while others may be deleteriously affected by
the compounded effect of both drivers [@deleo2017].~~
